---
title: "2026_02_02"
author: "Yuan Xie"
format: "html"
---

## **Change in place syntax**  

# Square bracket syntax
```{r}
z <- c(latte = 2, boba = 4, matcha = 6)
z[c(TRUE, FALSE, TRUE)] <- 12
z
# Show the names of the vector
names(z)
# Change the names of the vector
names(z) <- c("A", "B", "C")
z
```

## **Matrix**  
```{r}
mat1 <- matrix(data = c(1, 2, 3, 4), nrow = 2, ncol = 2) ## column-major
mat1
typeof(mat1)
```
Note: defaults to fill the matrix by columns (byrow = FALSE)

```{r}
mat1 <- matrix(data = c(1, 2, 3, 4), nrow = 2, ncol = 2)
mat1

mat2 <- matrix(data = c(1, 2, 3, 4), nrow = 2)
mat2

mat3 <- matrix(data = c(1, 2, 3, 4), ncol = 2)
mat3
## row-major here
mat4 <- matrix(data = c(1, 2, 3, 4), nrow = 2, byrow = TRUE)
mat4
mat5 <- matrix(data = c(1, 2, 3, 4), ncol = 2, byrow = TRUE)
mat5
```

# Matrices as vectors
```{r}
vec <- 1:4
vec
dim(vec) <- c(2, 2)
vec
```
Note: Matrices are simply vectors with dimensions dim(). All vector behavior still applies.

# Questions
```{r}
matrix(c(1, 2, 3, 4), nrow = 2) *
    matrix(c(1, 2, 3, 4), nrow = 2)

matrix(c(1, 2, 3, 4), nrow = 2) %*%
    matrix(c(1, 2, 3, 4), nrow = 2)

matrix(data = c("iz", "buzz", "poppy"), nrow = 2)

matrix(c(TRUE, 3L, 1.0, FALSE), nrow = 2)
```

```{r}
m1 <- matrix(1:4, nrow = 2)
m2 <- matrix(1:9, nrow = 3)
# m1 * m2 will give you an error
```

## Matrix subsetting  

# By index
```{r}
mat <- matrix(1:4, nrow = 2)
mat
mat[1, 2]

mat2 <- matrix(1:9, nrow = 3)
mat2
mat2[1:2, 2:3]

mat2 <- matrix(1:9, nrow = 3)
mat2
mat2[1, ]
mat2[1, , drop = FALSE]
mat2[-2, ]
```
Notes:  
1. subsetting a matrix row creates a vector.\
2. empty subsetting vectors > all entries.\

# By names
```{r}
mat <- matrix(1:4, nrow = 2)
rownames(mat) <- c("a", "b")
colnames(mat) <- c("i", "ii")
mat
mat["a", "ii"]
mat["a", ]
mat[ ,"ii"]
```

# By logical
```{r}
mat <- matrix(1:4, nrow = 2)
rownames(mat) <- c("a", "b")
colnames(mat) <- c("i", "ii")
mat[c(TRUE, FALSE), c(FALSE, TRUE)]
mat["a", c(FALSE, TRUE)]
```

```{r}
mat_log <- matrix(c(TRUE, FALSE, TRUE, FALSE), nrow = 2)
mat[mat_log]
```

## **Matrices and Tables**
```{r}
Name <- c("Leia", "Luke", "Han")
Height <- c(1.50, 1.70, 1.80)
Force <- c(TRUE, TRUE, FALSE)

M1 <- matrix(c(Name, Height, Force), nrow = 3)
M1
colnames(M1) <- c("Name", "Height", "Force")
M1
```

```{r}
Name <- c("Leia", "Luke", "Han")
Height <- c(1.50, 1.70, 1.80)
Force <- c(TRUE, TRUE, FALSE)

M2 <- cbind(Name, Height, Force) # column-bind
M2
```

```{r}
leia <- c("Leia", 1.50, TRUE)
luke <- c("Luke", 1.70, TRUE)
han <- c("Leia", 1.80, FALSE)

M3 <- rbind(leia, luke, han) # row-bind
colnames(M3) <- c("Name", "Height", "Force")
M3
```












