---
title: "2026_02_04"
author: "Yuan Xie"
format: 
    html:
        fontsize: 1.1em
        theme: darkly
embed-resources: true
---

## **Review**  

We learned atomic vectors with factors and matrix.\
1. Atomic vectors: 1-dim\
2. Matrix: 2-dim array; dim1 = rows, dim2 = cols\
3. Arrays: >= 3-dim; use array()\

# Question: Which of the following four objects are the same?
```{r}
seq(1, 6)

seq(1, 3) + seq(4, 6)

c(seq(1, 3), seq(4, 6))

c(c(seq(1, 3), seq(4, 6)))
```
Note:\
1. All elements of the same type\
2. Cannot contain other vectors\

## **Lists**  

A list is a non-atomic vector that can contain:\
1. Atomic vectors of different types\
2. Other lists\

```{r}
list(1, "two")
```
Note: [[1]] means the first element, and [[2]] means the second element.

```{r}
list(1:3, "a", 4:6)
```

# Subsetting with []  

[] extracts elements of a vector (atomic or non-atomic) using a subsetting (atomic) vector.\
1. Atomic vectors and matrices return atomic vectors.\
2. Lists return lists\

```{r}
vec <- 1:4
vec[1:2]

mat <- matrix(vec, nrow = 2)
mat
```

```{r}
x <- list(1:3, "a", 4:6)
x[1]
x[2:3]
```

# Subsetting with names
```{r}
named_x <- list("Ape" = 1:3, "Boy" = "a", "Cat" = 4:6)
named_x

named_x[c("Boy", "Cat")]
```

# Subsetting with logicals
```{r}
named_x <- list("Ape" = 1:3, "Boy" = "a", "Cat" = 4:6)
named_x[c(FALSE, TRUE, FALSE)]
```


## Question: What if you want the actual data structure inside the list?\
[[i]] extracts original object inside the list.\

```{r}
x <- list(1:3, "a", 4:6)
x[[1]]

# !length has difference
length(x[1])
length(x[[1]])
```

# Exercise
```{r}
x <- list(1:3, "a", 4:6)
x[1:2]
x[[3]]
x[-2]
x[c(1, 2)]
c(x[[3]], x[[1]])
```

# Subsetting single elements with $  
\
$ will extract a single element of a list by name.
```{r}
named_x <- list("Ape" = 1:3, "Boy" = "a", "Cat" = 4:6)
named_x$Cat
named_x$C ### Permits partial matching.
```

## Nesting lists 

A list is a non-atomic vector that can contain:\
1. Atomic vectors of different types\
2. Other lists\

```{r}
y <- list(1, "two", list(TRUE, 1:3))
y[2]
y[[2]]
y[[3]][[2]]
y[[3]][[2]][2]
```




































