---
title: "2026_01_28"
author: "Yuan Xie"
format: "html"
---

**Remark of vectors**  
1. Ordered sets of elements.  
2. Atomic: elements of the same type.  
3. Have length().  
4. Elements can be named: names().  
5. Types: logical, integers, double, character; Modes: logical, numeric, character.  

```{r}
x <- c("a" = 2, "b" = 4)
x
```

**Creating vectors**

Functions to **initialize** "dummy" vectors.

```{r}
logi <- logical(length = 2)
logi
```
Note: with **FALSE** default

```{r}
int <- integer(length = 3)
int

dbl <- double(length = 4)
dbl
```
Note: with **0** default

```{r}
char <- character(length = 5)
char
```
Note: with **""** default

**Functions to create vectors**

```{r}
logi <- vector(mode = "logical", length = 2)
logi
int <- vector(mode = "integer", length = 3)
int
dbl <- vector(mode = "double", length = 4)
dbl
char <- vector(mode = "character", length = 5)
char
```

**Numeric sequences with seq()**

```{r}
seq(from = 1, to = 5, by = 1)
seq(from = 1, to = 10, by = 2)
seq(from = 10, to = 1, by = -2)
seq(from = 1, by = 2, length.out = 6)
```

**Numeric sequences with :**

```{r}
1:5
1:10
-3:3
-3.5:3.5
```

**Replicates with rep()**

```{r}
rep(1, times = 3)
rep(1:3, times = 2)
rep(1:3, each = 2, times = 2)
rep(1:3, each = 2, times = 2, length.out = 10)
```

```{r}
########## Exercise
y <- c(0, 1, 2)
rep(y, times = y + 1)
rep(y, times = 1:3)
```

**Some special cases**

```{r}
c(1, 2, 3) + c(2, 4, 8)
c(1, 2, 3) + c(2, 4)
c(1, "dog", TRUE)
c(1, TRUE)
c("dog", TRUE)
```

**Vectorization**  

Vectorized: Operations and functions that can run on an entire vector at once.  
```{r}
dbl_vec <- c(1, 2.3, 4)

# Logarithm function log()
log(dbl_vec)

sum(dbl_vec)
```

**Vector Recycling**  

Vector Recycling: If a function requires a longer vector than provided, it will full the vector with the existing elements, in order
```{r}
c(2, 6, 1) ^ 2
c(2, 6, 3) ^ c(2, 3)
```

**Coercion**  

Implicit Coercion: Automatic change of data types to ensure atomic vectors.
```{r}
c(1, "dog", TRUE)
```
Changes follow a hierarchy: logical < integer < double < character  

Explicit Coercion: R provides functions to explicitly coerce from one type to another.  
as.logical(), as.integer(), as.double(), as.character()  
```{r}
as.double(c(TRUE, FALSE, FALSE, FALSE))
```


**Square Bracket Syntax**  

```{r}
x <- c(2, 4, 6)
# By index
x[2]
```

**Subsetting by Index**  

```{r}
char_vec <- c("bear", "owl", "squirrel")
char_vec
char_vec[2]
```










