---
title: "2026_01_30"
author: "Yuan Xie"
format: "html"
---

## **Subsetting with vectors**  

# Square bracket syntax
```{r}
vec <- c("a" = 1, "b" = 2, "c" = 3)
# vec[indexing vector, numeric, char, logical]
```

```{r}
char_vec <- c("bear", "owl", "squirrel")
char_vec
char_vec[2]
char_vec[2:3]
char_vec[-2]
char_vec[-3:-2]
```

# Subset by name  
```{r}
char_vec2 <- c("oski" = "bear", "fiat" = "owl","lux" = "squirrel")
char_vec2
char_vec2["oski"]
char_vec2[c("lux", "fiat")]
```

# Subset by logicals  
```{r}
char_vec2 <- c("oski" = "bear", "fiat" = "owl","lux" = "squirrel")
char_vec2
char_vec2[c(TRUE, FALSE, TRUE)]

# You cannot type only one true. The length should be equal. Otherwise, recycling wll take place.
char_vec2[TRUE]
```

# Exercise 1
```{r}
z <- c(latte = 2, boba = 4, matcha = 6)

# Vector 1
z[c(2, 1)]
z[2:1]
z[c("boba", "latte")]

# Vector 2
z[c(3, 3)]
z[rep(3, times = 2)]
z[c("matcha", "matcha")]
```

# Exercise 2
```{r}
x <- c(2, 4, 6)
y <- c("a", "b", "c")

y[x[-3][2]]
y[x/x]
y[as.logical(x)]
```

## Change in place syntax  
```{r}
z <- c(latte = 2, boba = 4, matcha = 6)
z[c(TRUE, FALSE, TRUE)] <- 12
z

names(z)
names(z) <- c("A", "B", "C")
z
```

## Basic Unix  
```{bash eval = FALSE}
quarto render demo.qmd
quarto preview demo.qmd
quarto publish demo.qmd
```

# GUIs and CLIs
Windows and Mac use a Graphical User Interface (GUI) for you to interact with the Operating System.  
\
About GUIs: (See slides)
About CLIs: (See slides)
\
User -> Terminal -> Shell -> Operating System

# Shell commands
```{bash eval = FALSE}
pwd
ls
cd
mkdir
```

# Anatomy of Shell Commands




















