---
title: "2026_02_11"
author: "Yuan Xie"
format: 
    html:
        fontsize: 1.1em
        theme: darkly
embed-resources: true
---

# Factors\
```{r}
uid <- c("3850278", "5869204", "4728112", "3829651", "3859278")
maj <- c("STAT", "ECON", "STAT", "DATA", "ECON")
```
R does not have dictionary or hash table structures.\
Factors: A data structure for storing **categorical** data.\
Created with ```factor()```\
Stores the values in an integer vector and adds a levels attribute to map integers to character strings.\

```{r}
maj <- factor(x = c("STAT", "ECON", "STAT", "DATA", "ECON"))
maj
```

```{r}
as.integer(maj)
```

```{r}
class(maj)
```

## Attributes\
Objects built on top of simpler data structures can have attributes. A factor’s attributes include its class and levels.\
```{r}
attributes(maj)
```

```{r}
levels(maj)
```

## Treat factors like vectors\
```{r}
maj
length(maj)
maj[1]
maj[-2]
```

```{r}
maj
levels(maj) <- c("DATA", "ECON", "STATISTICS")
maj
```

Add a level even if it’s **not observed**.\
```{r}
levels(maj) <- c("DATA", "ECON", "STATISTICS", "BIO")
maj
```

## Taking advantage of class\
```{r}
class(1:4)
summary(1:4)
class(maj)
summary(maj)
```

## Ordering levels
```{r}
maj_or <- factor(x = c("STAT", "ECON", "STAT", "DATA", "ECON"),
                 levels = c("STAT", "ECON", "DATA", "BIO"),
                 ordered = TRUE)
maj_or

sort(maj_or)
```

```{r}
barplot(summary(maj_or))
barplot(summary(maj))
```

## Factors and Matrices
```{r}
mat <- matrix(1:4, ncol = 2)
class(mat)

summary(mat)
```

# Data Frame\
A data frame is a named **list** of vectors of the same length. Created with ```data.frame()```.
```{r}
star_wars <- data.frame(
  name = c("Anakin", "Padme", "Luke", "Leia"),
  gender = c("male", "female", "male", "female"),
  height = c(1.88, 1.65, 1.70, 1.50),
  weight = c(84, 45, 77, 49)
)

star_wars
```

```{r}
attributes(star_wars)
```

```{r}
names(star_wars)
row.names(star_wars)
class(star_wars)
```

## Other useful functions for Data Frames\
```{r}
library(tidyverse)
```


```colnames()```: equivalent to ```names()```\
```rownames()```: equivalent to ```row.names()```\
```summary()```: summary of each column\
```str()```: structure\
```head()```: first n rows\
```tail()```: last n rows\
```dim()```: dimensions\
```ncol()```: number of columns\
```nrow()```: number of rows\
```{r}
colnames(star_wars)
rownames(star_wars)
summary(star_wars)
str(star_wars)

head(star_wars)
star_wars %>% head()

tail(star_wars)
star_wars %>% tail()

dim(star_wars)

ncol(star_wars)
nrow(star_wars)
```















