---
title: "2026_02_09"
author: "Yuan Xie"
format: 
    html:
        fontsize: 1.1em
        theme: darkly
embed-resources: true
---

# Functions
```{r}
expr <- {x <- 1
         y <- 2
         z <- c(x, y)}
expr
```

```{r}
# a and b: global objects
a <- 10
b <- 20

hey <- function(){
  # a and b: local objects
  a <- 1
  b <- 2
  c(a, b)
}

hey() # execution, ephemeral
```

## Ex. colSums()
```{r}
mat <- matrix(1:4, nrow = 2)
mat
colSums(mat)
```
What if there is no colSums()?
```{r}
colsum_vec <- rep(0, ncol(mat))
for (coln in 1:ncol(mat)) {
  colsum_vec[coln] <- sum(mat[, coln])
}

colsum_vec
```

```{r}
# celsius to fahrenheit
# one or more arguments
c2f <- function(celsius){
  (9/5) * celsius + 32
}

c2f(0)
c2f(1:10)
c2f(c(2, 2, NA))
```

3 aspects: \
1. arguments\
2. body\
3. creation environment\ (different from execution environment)

```{r}
a <- 2 # global
yay <- function(){
  b <- 1 # internal
  c(a, b)
}

yay()
```
The environment within the function includes the environment “one level up”\
Objects defined in the function environment mask those one level up\
Best to encapsulate functions (not make them dependent on objects defined outside their environment / arguments)\

```{r}
roll_die <- function(){
  sample(x = 1:6, size = 1)
}
```

work for arbitrary n-sided dice, return more than 1 roll of the die
```{r}
roll_die <- function(n_sides = 6, n_rolls = 1){
  sample(x = 1:n_sides, size = n_rolls, replace = TRUE)
}
```

## stop() and warning()
```{r}
set.seed(133)

roll_die <- function(n_sides = 6, n_rolls = 1){
  if (n_sides <= 0){
    stop("invalid value n-sides")
  }
  if (n_rolls <= 0){
    stop("invalid value n-rolls")
  }
  sample(x = 1:n_sides, size = n_rolls, replace = TRUE)
}

roll_die(6, 10)
```






















