---
title: "Crash Data 2023"
subtitle: "STAT 133 - Project 2"
author: "Yuan Xie (3038930130)"
format:
  revealjs:
    theme: simple
    code-fold: true
    slide-number: c
    smaller: true
    transition: fade
    embed-resources: true
    title-slide-attributes:
      data-background-color: pink
---

# 2.1 Main content (about slide) {background-color="indigo" transition="fade" transition-speed="slow"}

##  {background-color="indigo" transition="fade" transition-speed="slow"}
This analysis project is about the crash data in California during 2023 from the Statewide Integrated Traffic Records System (SWITRS). The data set is collected and
maintained by the California Highway Patrol (CHP).\

We will examine:\

:::{.incremental}
- Overall crash statistics
- Distribution and ranking tables
- Crashes involving pedestrians and bicyclists
- Exploratory research questions
:::

## Data Import and Briefly Show the Data Structure {background-color="indigo" transition="fade" transition-speed="slow"}
```{r}
#| echo: true
library(tidyverse)
# use this code chunk to import your data
dat <- read_csv("crashes_2023.csv")
head(dat)
```

# 2.2 Overall Statistics {background-color="cyan" transition="fade" transition-speed="slow"}

## Total Number of Crashes, Injured Victims and Fatal Victims {background-color="cyan" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
overall_stats <- dat |>
  summarise(
    total_crashes = n(),
    total_injured = sum(NUMBER_INJURED, na.rm = TRUE),
    total_killed  = sum(NUMBER_KILLED, na.rm = TRUE)
  )

knitr::kable(overall_stats, caption = "Overall Statistics Data")
```

In 2023, there were **`r overall_stats$total_crashes`** crashes in California. These crashes resulted in **`r format(overall_stats$total_injured, scientific = FALSE)`** injured victims and  **`r overall_stats$total_killed`** fatal victims.

# 2.3 Distribution Tables {background-color="aquamarine" transition="fade" transition-speed="slow"}

## Counts and Percentages Per Month {background-color="aquamarine" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
month_dist <- dat |>
  group_by(MONTH) |>
  summarise(count = n()) |>
  mutate(percent = round(count / sum(count) * 100, 2)) |>
  arrange(MONTH)

left  <- month_dist[1:6, ]
right <- month_dist[7:12, ]
```

:::: columns

::: column
```{r}
#| echo: true
knitr::kable(left)
```
:::

::: column
```{r}
#| echo: true
knitr::kable(right)
```
:::

::::

## Counts and Percentages Per Day of Week {background-color="aquamarine" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
day_dist <- dat |>
  group_by(DAY_OF_WEEK) |>
  summarise(count = n()) |>
  mutate(percent = round(count / sum(count) * 100, 2)) |>
  arrange(DAY_OF_WEEK)
knitr::kable(day_dist)
```

## Counts and Percentages Per Day of Hour {background-color="aquamarine" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
hour_dist <- dat |>
  group_by(HOUR) |>
  summarise(count = n()) |>
  mutate(percent = round(count / sum(count) * 100, 2)) |>
  arrange(HOUR)

h1 <- hour_dist[1:6, ]
h2 <- hour_dist[7:12, ]
h3 <- hour_dist[13:18, ]
h4 <- hour_dist[19:24, ]
```

See tables in the next slide.

##  {background-color="aquamarine" transition="fade" transition-speed="slow"}

::: {.smaller}

:::: columns

::: column
```{r}
knitr::kable(h1)
```
:::

::: column
```{r}
knitr::kable(h2)
```
:::

::: column
```{r}
knitr::kable(h3)
```
:::

::: column
```{r}
knitr::kable(h4)
```
:::

::::

:::

# 2.4 Ranking Tables {background-color="orange" transition="fade" transition-speed="slow"}

##  Top-10 PCF Violations {background-color="orange" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top10_pcf <- dat |>
  group_by(PCF_VIOL_CATEGORY) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:10)

knitr::kable(top10_pcf)
```

##  Types of Collision Rank {background-color="orange" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
collision_rank <- dat |>
  group_by(TYPE_OF_COLLISION) |>
  summarise(count = n()) |>
  arrange(desc(count))

knitr::kable(collision_rank)
```

## Top-8 Party At Fault {background-color="orange" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top8_fault <- dat |>
  group_by(STWD_VEHTYPE_AT_FAULT) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:8)

knitr::kable(top8_fault)
```

# 2.5 Crashes Involving Pedestrians {background-color="grey" transition="fade" transition-speed="slow"}

## Number of Accidents Involving Pedestrians, Injured Pedestrians and Killed Pedestrians {background-color="grey" transition="fade" transition-speed="slow"}
```{r}
#| echo: true
ped_accident <- dat |>
  filter(PEDESTRIAN_ACCIDENT == "yes") |>
  summarise(
    total_ped_crashes = n(),
    total_ped_injured = sum(COUNT_PED_INJURED, na.rm = TRUE),
    total_ped_killed  = sum(COUNT_PED_KILLED, na.rm = TRUE)
  )

knitr::kable(ped_accident, caption = "Pedestrian Statistics Data")
```

There were **`r ped_accident$total_ped_crashes`** crashes involving pedestrians. The crashes resulted in **`r format(ped_accident$total_ped_injured, scientific = FALSE)`** injured pedestrians and **`r ped_accident$total_ped_killed`** pedestrian fatalities.

##  Top-5 PCF Violations for Pedestrian Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_pcf_ped <- dat |>
  filter(PEDESTRIAN_ACCIDENT == "yes") |>
  group_by(PCF_VIOL_CATEGORY) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_pcf_ped)
```

## Top-5 Types of Collision for Pedestrian Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_collision_ped <- dat |>
  filter(PEDESTRIAN_ACCIDENT == "yes") |>
  group_by(TYPE_OF_COLLISION) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_collision_ped)
```

## Top-5 Party at Fault for Pedestrian Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_fault_ped <- dat |>
  filter(PEDESTRIAN_ACCIDENT == "yes") |>
  group_by(STWD_VEHTYPE_AT_FAULT) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_fault_ped)
```

# 2.6 Crashes Involving Bicyclists {background-color="grey" transition="fade" transition-speed="slow"}

## Number of Accidents Involving Bicyclists, Injured Bicyclists and Killed Bicyclists {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
bic_accident <- dat |>
   filter(BICYCLE_ACCIDENT == "yes") |>
  summarise(
    total_bic_crashes = n(),
    total_bic_injured = sum(COUNT_BICYCLIST_INJURED, na.rm = TRUE),
    total_bic_killed  = sum(COUNT_BICYCLIST_KILLED, na.rm = TRUE)
  )

knitr::kable(bic_accident, caption = "Bicyclist Statistics Data")
```

There were **`r bic_accident$total_bic_crashes`** crashes involving bicyclists. The crashes resulted in **`r format(bic_accident$total_bic_injured, scientific = FALSE)`** injured bicyclists and **`r bic_accident$total_bic_killed`** bicyclist fatalities.

##  Top-5 PCF Violations for Bicyclist Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_pcf_bic <- dat |>
  filter(BICYCLE_ACCIDENT == "yes") |>
  group_by(PCF_VIOL_CATEGORY) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_pcf_bic)
```

## Top-5 Types of Collision for Bicyclist Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_collision_bic <- dat |>
  filter(BICYCLE_ACCIDENT == "yes") |>
  group_by(TYPE_OF_COLLISION) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_collision_bic)
```

## Top-5 Party at Fault for Bicyclist Accidents {background-color="grey" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
top5_fault_bic <- dat |>
  filter(BICYCLE_ACCIDENT == "yes") |>
  group_by(STWD_VEHTYPE_AT_FAULT) |>
  summarise(count = n()) |>
  arrange(desc(count)) |>
  slice(1:5)

knitr::kable(top5_fault_bic)
```

# 2.7 Exploratory Research Question {background-color="violet" transition="fade" transition-speed="slow"}

## Question: Is there a relationship between hour of the day and number of crashes? {background-color="violet" transition="fade" transition-speed="slow"}

```{r}
#| echo: true
hour_summary <- dat |>
  group_by(HOUR) |>
  summarise(crashes = n())

ggplot(hour_summary, aes(x = HOUR, y = crashes)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Number of Crashes by Hour of Day",
    x = "Hour of Day",
    y = "Number of Crashes"
  )
```

## Answer {background-color="violet" transition="fade" transition-speed="slow"}

Based on our data analysis and the bar chart plotted, it seems that there is some relationship between hour of the day and number of crashes. We can observe that:\

:::{.incremental}
- The number of crashes is the lowest during early morning hours, especially 3am and 4am.
- There is an substantial increase in crashes during the morning commute, especially 7am and 8 am.
- Most crashes happen in the late afternoon, around 4pm and 5pm.
- The number of crashes decreases gradually after 6pm.
:::

# Thanks for your reading and patience! {background-color="beige" transition="fade" transition-speed="slow"}























