---
title: "Project 3"
author: "Yuan Xie (3038930130)"
format: 
  closeread-html:
    theme: cosmo
    cr-style:
      section-background-color: white
      narrative-background-color-sidebar: black
      narrative-text-color-sidebar: white
      narrative-font-size: 2em
embed-resources: true
---

```{r setup, include=FALSE}
# load packages and/or import data set(s)
library(tidyverse)
nba <- read_csv("nba2025-clean.csv")
```

:::: {.cr-section}

The main goal in this project is to craft a story around salaries of NBA players and payrolls of NBA teams in season 2024-2025.

@cr-a1

1) We are going to draw one graph to visualize the salary distribution of all players. We divide the salary by 1000000 to make the graph look easier.

::: {#cr-a1}

```{r}
ggplot(nba, aes(x = salary/1000000)) +
  geom_histogram(bins = 30, fill = "orange", color = "white") +
  labs(
    title = "Distribution of NBA Player Salaries (2025)",
    x = "Salary (USD)",
    y = "Number of Players"
  ) + 
  theme_minimal()
```

:::

[@cr-a1]

@cr-b1

2) This is the table with summary statistics of salaries among all players.

::: {#cr-b1}

```{r}
salary_summary <- summary(nba$salary)

knitr::kable(
  data.frame(
    Statistic = names(salary_summary),
    Value = as.numeric(salary_summary)
  )
)
```

:::

[@cr-b1]{pan-to="-100px, 0px"}

@cr-c1

3) This is a table showing  the top-10 highest-paid players (name, team, position, salary).

::: {#cr-c1}

```{r}
c1 <- nba |>
  arrange(desc(salary)) |>
  slice(1 : 10) |>
  select(player, team, position, salary)
knitr::kable(c1)
```

:::

[@cr-c1]{scale-by="2"}

@cr-d1

4) This is a graph that explores the relationship between salary and experience. Since the slope is positive, there may be a positive relationship between salary and experience.

:::{#cr-d1}

```{r}
nba |>
  ggplot(aes(x = experience, y = salary/1000000)) +
  geom_point(alpha = 0.6, aes(color = position)) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, color = "red") +
  labs(
    title = "Relationship Between Player Experience and Salary",
    x = "Years of Experience",
    y = "Salary (Million USD)"
  ) +
  theme_minimal()
```

:::

[@cr-d1]{pan-to="0px, 100px"}

@cr-e1

5) This is a graph that ranks the teams by salaries.

:::{#cr-e1}

```{r}
nba_team_salary <- nba |>
  group_by(team) |>
  summarise(payroll = sum(salary, na.rm = TRUE))

nba_team_salary |>
  ggplot(aes(x = reorder(team, payroll), y = payroll/1000000)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "NBA Team Payroll Rankings",
    x = "Team",
    y = "Total Payroll (Million USD)"
  ) +
  theme_minimal()
```

:::

[@cr-e1]{pan-to="0px, 100px"}

@cr-f1

6) My own exploratory analysis: Salary Distribution Across Player Positions

:::{#cr-f1}

```{r}
nba |>
  ggplot(aes(x = position, y = salary/1000000)) +
  geom_boxplot(fill = "steelblue", alpha = 0.7) +
  labs(
    title = "Salary Distribution Across Player Positions",
    x = "Position",
    y = "Salary (Million USD)"
  ) +
  theme_minimal()
```

:::

Observation: PG has the greatest salary spread while the rest four positions have outliers. However, the median salary is relatively similar across positions.

[@cr-f1]

::::








