---
title: "PS1: Vectors"
subtitle: "STAT 133"
author: "Yuan Xie"
format: 
  html:
    toc: true
    number-sections: true
    theme: simplex
embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


::: {.callout-note icon=false}

## General Instructions

-   Use the template file `ps1.qmd` to answer the questions.

-   Rename this file as `ps1-first-last.qmd`, where `first` and `last`
    are your first and last names (e.g. `ps1-gaston-sanchez.qmd`).

-   Submit both your `qmd` and `HTML` files to the corresponding assignment
    submission in bCourses.

:::


# Numeric Sequences

Use the `seq()` function to create the following vectors:

a) `1 2 3 4 5 6 7 8 . . . 18 19 20`

```{r}
# your code
seq(from = 1, to  = 20)
```

\

b) `1.0 1.2 1.4 1.6 1.8 2.0`

```{r}
# your code
seq(from = 1.0, to  = 2.0, by = 0.2)
```

\

c) `10 8 6 4 2 0 -2 -4 -6 -8 -10`

```{r}
# your code
seq(from = 10, to = -10, by = -2)
```

\

d) A sequence with the first 20 multiples of 3: `3 6 9 12 ...`. _Hint_: the `length.out` argument in `seq()` is your friend.

```{r}
# your code
seq(from = 3, by = 3, length.out = 20)
```


-----


# Replicates

The function `rep()` allows you to create vectors of replicated elements in a handful of different ways. The main input is a vector `x`, and you can specify arguments such as `each`, `times` and `length.out`

```r
rep(x, each, times, length.out)
```


a) Figure out how to use the `rep()` function and the input vector `1:3` to create the output `1 1 2 2 3 3`

```{r}
# your code
vec_2a <- c(1:3)
rep(vec_2a, each = 2)
```

\

b) Figure out how to use the `rep()` function and the input vector `1:3` to create the output `1 2 3 1 2 3`

```{r}
# your code
vec_2b <- c(1:3)
rep(vec_2b, times = 2)
```

\

c) Figure out how to use the `rep()` function and the input vector `1:3` to create the output `1 1 2 2 3 3 1 1 2 2 3 3`

```{r}
# your code
vec_2c <- c(1:3)
rep(1:3, each = 2, times = 2)
```

\

d) Figure out how to use the `rep()` function and the input vector `1:4` to create the output `1 2 2 3 3 3 4 4 4 4`

```{r}
# your code
vec_2d <- c(1:4)
rep(vec_2d, times = 1:4)
### I have another way
rep(vec_2d, times = vec_2d)
```


-----


# GSW Data (season 2025)

The (raw) data for this problem has to do with some players from Golden State 
Warriors (2025), displayed in the table below (source: Basketball-Reference).

| number | player             | position | height | weight | experience |
|--------|--------------------|----------|--------|--------|------------|
|   30   | Stepehn Curry      |  PG      |  74    |  185   |    16      |
|   10   | Jimmy Buttler      |  SF      |  78    |  230   |    14      |
|   23   | Draymond Green     |  PF      |  78    |  230   |    13      |
|    2   | Brandin Podziemski |  SG      |  76    |  205   |     2      |
|    1   | Jonathan Kuminga   |  PF      |  79    |  225   |     4      |
|    4   | Moses Moody        |  SG      |  77    |  211   |     4      |
|   20   | Al Horford         |   C      |  80    |  240   |    18      |
|    3   | Will Richard       |  SG      |  75    |  206   |     0      |


## Creating Vectors (by hand)

Use the "combine" function `c()` to create vectors for the columns in the data table displayed above, 
according to the following data types.

- `number`: integer vector

- `player`: character vector

- `position`: character vector

- `height`: double (i.e. real) vector

- `weight`: double (i.e. real) vector

- `experience`: integer vector


```{r}
# your code
number <- c(30L, 10L, 23L, 2L, 1L, 4L, 20L, 3L)
number
typeof(number)

player <- c("Stepehn Curry", "Jimmy Buttler", "Draymond Green", "Brandin Podziemski", "Jonathan Kuminga",
            "Moses Moody", "Al Horford", "Will Richard")
player
typeof(player)

position <- c("PG", "SF", "PF", "SG", "PF", "SG", "C", "SG")
position
typeof(position)

height <- c(74, 78, 78, 76, 79, 77, 80, 75)
height
typeof(height)

weight <- c(185, 230, 230, 205, 225, 211, 240, 206)
weight
typeof(weight)

experience <- c(16L, 14L, 13L, 2L, 4L, 4L, 18L, 0L)
experience
typeof(experience)
```


## Subsetting (i.e. indexing) Vectors

Refer to the vectors created in the preceding subsection.
Write R commands (just one single-line commands)---displaying the output---that
answer the following questions:


a) What is the weight of the heaviest player?

```{r}
# your code
max(weight) # The heaviest player is the 7th one.
```

<br>

b) What is the name of the player that has a height of 80 inches`?

```{r}
# your code
player[height == 80] # Al Horford has a height of 80 inches.
```

<br>

c) What is the position of the player with more years of experience?
_Hint_: the `which.max()` function is your friend.

```{r}
# your code
position[which.max(experience)]
```

<br>

d) What is the number of the lightest player? _Hint_: the `which.min()` 
function is your friend.

```{r}
# your code
number[which.min(weight)]
```

<br>

e) What is the average weight for those players with more than 5 years of
experience?

```{r}
# your code
# use mean() function to calculate the average
mean(weight[experience > 5])
```

<br>

f) How many players have a weight larger than the average (i.e. mean) weight?
_Hint_: the `sum()` function is your friend.

```{r}
# your code
sum(weight > mean(weight))
```

<br>

g) How many players have between 9 and 12 years of experience (inclusive)?

```{r}
# your code
# & means and in R language
sum(experience >= 9 & experience <= 12)
```

<br>

h) What is the median years of experience of _Shooting Guard_ (`SG`) players?

```{r}
# your code
median(experience[position == "SG"])
```

<br>

i) What is the median weight of those players with a position different of _Center_ (`C`)?

```{r}
# your code
median(weight[position != "C"])
```

<br>

j) What is the first quartile (i.e. bottom 25th percentile) of years of
experience among _Power Forwards_ (`PF`) and _Shooting Guards_ (`SG`).
_Hint_: the `quantile()` function is your friend.

```{r}
# your code
quantile(experience[position == "PF" | position == "SG"])["25%"] # subsetting by name
```


-----


# Subsetting Commands

Consider the vector `vec = c("x", "y", "a", "b")`. For each of the following 
parts, write at least two different subsetting commands that produce the given 
outputs:


a) `"x"`

```{r}
# your code
vec = c("x", "y", "a", "b")

vec[1] # 1st way, square bracket syntax
vec[c(TRUE, FALSE, FALSE, FALSE)] # 2nd way, subsetting by logicals
vec[c(-2, -3, -4)] # 3rd way, remove the the other 3 letters
vec[-4:-2] # 4th way, remove the the other 3 letters
```

\

b) `"b" "a" "y" "x"`

```{r}
# your code
vec = c("x", "y", "a", "b")

vec[c(4, 3, 2, 1)] # 1st way, square bracket syntax with a vector
vec[4:1] # 2nd way, square bracket syntax with sequence
```

\

c) `"b" "b" "b"`

```{r}
# your code
vec = c("x", "y", "a", "b")

vec[c(4, 4, 4)] # 1st way, square bracket syntax, using a vector to repeat
vec[rep(4, times = 3)]  # 2nd way, square bracket syntax with repetition
```

\

d) `"x" "y`

```{r}
# your code
vec = c("x", "y", "a", "b")

vec[c(1, 2)] # 1st way, square bracket syntax with a vector
vec[1:2]  # 2nd way, square bracket syntax with sequence
vec[c(-3, -4)] # 3rd way, remove the "a" and "b"
vec[-3:-4] # 4th way, remove the "a" and "b"
```


-----


# Technical questions about vectors

Consider the following two vectors `x` and `y`:

```{r}
x <- c(1, 2, 3, 4, 5)

y <- c("a", "b", "c", "d", "e")
```

In the following parts, indicate True or False for whether the given command
involves: a) coercion, and/or b) vectorization, and/or c) recycling.


1) `y[x <= 2L]`

    a) coercion: True. Because the integer is changed to double.
    b) vectorization: True. Because it makes a comparison through the whole vector x.
    c) recycling: True. Because length(x) = 5 and length(2L) = 1.


2) `y[x[5]]`

    a) coercion: False. Because it is just a subset by index.
    b) vectorization: False. Because it is just a subset by index.
    c) recycling: False. Because it does not require a longer vector.


3) `x[y != 'c']`

    a) coercion: False. Because it is just subsetting by logicals.
    b) vectorization: True. Because it needs to check whether the character is c one by one.
    c) recycling: True. Because length(y) = 5 and length("c") = 1.


4) `y[x/x]`

    a) coercion: False. Because there is no change of data types.
    b) vectorization: True. Because it calculates x/x one by one although the results are all 1.
    c) recycling: False. Because the length of vectors are the same.


-----


# File Paths

Consider the following file structure depicted below. Note that the first forward slash `/` indicates the root directory.

```
/
  project/
    README.md
    data/
      rawdata/
        rawdata.csv
        rawdata-dictionary.md
      cleandata/
        cleandata.csv
        cleandata-dictionary.md
    code/
      functions.R
      clean-data-script.R
```

\

a) What is the absolute path to the folder `cleandata/`?

```{bash eval = FALSE}
/project/data/cleandata/
```

\

b) What is the relative path from `functions.R` to `cleandata.csv`?

```{bash eval = FALSE}
../data/cleandata/cleandata.csv
```

\

c) What is the relative path from `rawdata.csv` to `clean-data-script.R`?

```{bash eval = FALSE}
../../code/clean-data-script.R
```

\

d) Assume that your current directory is the `cleandata/` directory. Write the Unix command that lists the contents in the `project/` directory.

```{bash eval = FALSE}
ls ../..
```

\

e) Assume that your current directory is the `data/` directory. Write the Unix command that makes a new directory called `output/` located at the same level of `data/` and `code/`.

```{bash eval = FALSE}
mkdir ../output
```

